---
title: "dply Challenges with Solutions"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
#install.packages('dplyr')
library('dplyr')

gapminder <- read.csv("data/gapminder_data.csv", stringsAsFactors = TRUE)
```

## dplyr challenges

In breakout rooms of 3-4, work through these problems as a team. Choose one person to share their screen and the others in the room will help them solve the following challenges.   

### Challenge 1 
Write a single command (which can span multiple lines and includes pipes) that will produce a dataframe that has the African [or any other country you pick] values for lifeExp, country and year, but not for the other Continents. How many rows does your data frame have, and why?

```{r}
gapminder %>%
  filter(continent == "Africa") %>%
  select(year, country, lifeExp) %>%
  head
```


### Challenge 2  
Calculate the average life expectancy per country.
Which has the longest average life expectancy and which has the shortest average life expectancy?  

```{r}
lifeExp_bycountry <- gapminder %>%
  group_by(country) %>%
  summarize(mean_lifeExp = mean(lifeExp))

lifeExp_bycountry %>%
  filter(mean_lifeExp == min(mean_lifeExp) | mean_lifeExp == max(mean_lifeExp))
```

### Challenge 3  
Create a new dataframe that has a new column with the ratio of life expectancy to GDP per capita. Keep only the country and ratio column. 
**Hint:** think about the order of operations!   

```{r}
lifeExp_to_gdpPercap <- gapminder %>%
  mutate(lifeExp_ratio = lifeExp/gdpPercap) %>%
  select(country, lifeExp_ratio)

head(lifeExp_to_gdpPercap)
```

**2nd answer:**

```{r}
gapminder %>%
  group_by(country) %>%
  summarize(lifeExp_ratio = lifeExp/gdpPercap) %>%
  head()
```

### Challenge 4  
Calculate the average life expectancy in 2002 of 2 randomly selected countries for each continent. Then arrange the continent names in reverse order. 
Hint: Use the dplyr functions arrange() and sample_n(), they have similar syntax to other dplyr functions.

*Note*  sample_n() samples __n__ rows from a table. 

```{r}
gapminder %>% 
  filter(year==2002) %>%
  group_by(continent) %>%
  sample_n(2) %>%
  summarize(mean_lifeExp = mean(lifeExp)) %>%
  arrange(desc(mean_lifeExp))
```


### Challenge 5 
something with `filter()` and `count()`

Filter for continent (your choice) AND (&) a life expectancy greater then 60
Then count the observations for each group.

```{r}
gapminder %>%
  filter(continent == "Europe" & lifeExp > 60) %>%
  count(country) %>%
  head
  
```

